\documentstyle[a4, 12pt]{article}

\setlength{\parindent}{0ex}
\setlength{\parskip}{2.5ex}
\setlength{\textwidth}{14.8cm}
\pagestyle{empty}
\begin{document}

\begin{center}
{\large\bf Percentile rank gap as a measure of dependence across percentiles of a copula}\\

\vspace{0.5cm}
\underline{Weihao Choo$^1$}\\

\vspace{0.5cm}
{\it  $^1$Macquarie University, Sydney}
\end{center}

\hrulefill

{\bf Abstract}

This paper proposes ``percentile rank gap'' to quantify the dependence at various percentiles of a copula. The quantification is not always straightforward, but is important for example in stock markets where normal returns are weakly dependent but abnormal returns are highly dependent and linked to a major correction. Another example is insurance where natural catastrophe losses from related business lines happen simultaneously but typical losses are weakly dependent.

Percentile rank gap is defined in terms of first order conditional expectations, and lies between $-1$ and $1$ at all percentiles in the unit interval. For example, the percentile rank gap of a Gumbel copula starts below $1$ then increases to $1$, reflecting imperfect lower tail dependence but perfect upper tail dependence. For a Clayton copula, percentile rank gap starts at $1$, then decreases. In general percentile rank gap of $1$ at percentile $\alpha$ implies variables are simultaneously below or above percentile $\alpha$, or ``concordant'' relative to $\alpha$. Increasing ``discordance'' and dispersion between discordant percentiles reduces percentile rank gap.

Percentile rank gap satisfies several ``coherence'' properties. Percentile rank gap lies between $-1$ and $1$, with $-1$, $0$ and $1$ representing countermonotonicity, independence and comonotonicity, respectively. In addition percentile rank gap increases with the correlation order of the copula, therefore positively dependent variables have positive percentile rank gap, and vice versa. Lastly, taking a weighted average of percentile rank gap across all percentiles yields Spearman's correlation.

Percentile rank gap yields an alternative, practical approach to fit copulas to real world data. The starting point is empirically calculated percentile rank gap. The empirical calculation is refined by expert judgement of the dependence at various percentiles. Lastly a copula yielding the refined percentile rank gap is constructed. The resulting copula is subject to less uncertainty than the empirical copula, and is more tailored to actual data than parametric copulas.

\medskip




{\bf Keywords}
local dependence, copula, Spearman correlation.


\end{document}
